---
title: "DSII_homework2"
author: "Chirag Shah"
date: '2019-03-19'
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(caret)
library(glmnet)
library(splines)
library(ISLR)
library(boot)
library(mgcv)
```

Data Import 

```{r}
concrete = read_csv("concrete.csv") %>%
  janitor::clean_names()
```

#Part A

```{r, fig.height=5}
# matrix of predictors 
x <- model.matrix(compressive_strength~.,concrete)[,-1]
# vector of response
y <- concrete$compressive_strength
theme1 <- trellis.par.get()
theme1$plot.symbol$col <- rgb(.2, .4, .2, .5)
theme1$plot.symbol$pch <- 16
theme1$plot.line$col <- rgb(.8, .1, .1, 1)
theme1$plot.line$lwd <- 2
theme1$strip.background$col <- rgb(.0, .2, .6, .2)
trellis.par.set(theme1)
featurePlot(x, y, plot = "scatter", labels = c("","compressive strength"),
            type = c("p"), layout = c(4, 2))
```

#Part B

```{r}
#polynomial regression
fit1 <- lm(compressive_strength~water, data = concrete)
fit2 <- lm(compressive_strength~poly(water,2), data = concrete) 
fit3 <- lm(compressive_strength~poly(water,3), data = concrete) 
fit4 <- lm(compressive_strength~poly(water,4), data = concrete) 
```

```{r}
#cross-validation approach to find the best exponent
set.seed(1)

mseK <- rep(NA, 4)
for (i in 1:4) {
    fit <- glm(compressive_strength ~ poly(water, i), data = concrete)
    mseK[i] <- cv.glm(concrete, fit, K = 10)$delta[1]
}

plot(1:4, mseK, xlab = "Exponent", ylab = "Test MSE", type = "l")
```

The cross validation approach led me to believe that the appropriate polynomial function is of order 4. Lets confirm using the ANOVA test. 

```{r}
anova(fit1,fit2,fit3,fit4) 

plot1 <- ggplot(data = concrete, aes(x = water, y = compressive_strength)) +
     geom_point(color = rgb(.2, .4, .2, .5))
plot1
```

Using the anova test, models with an exponent of 2, 3, and 4 are significantly different from each other. However the results from the ANOVA doesn't tell you which is the optimal model (i.e. which polynomial function has the least mean squared error), only that they are different from the null. Therefore the type of results gained from the ANOVA test and cross-validation method are fundamentally different as cross validation is better for model selection in this case. 

```{r}
plot(compressive_strength ~ water, data = concrete, col = "green")
waterlims <- range(concrete$water)
water.grid <- seq(from = waterlims[1], to = waterlims[2], by = 1)

preds1 <- predict(fit1, newdata = data.frame(water = water.grid))
preds2 <- predict(fit2, newdata = data.frame(water = water.grid))
preds3 <- predict(fit3, newdata = data.frame(water = water.grid))
preds4 <- predict(fit4, newdata = data.frame(water = water.grid))

lines(water.grid, preds1, col = "red", lwd = 2)
lines(water.grid, preds2, col = "grey", lwd = 2)
lines(water.grid, preds3, col = "black", lwd = 2)
lines(water.grid, preds4, col = "purple", lwd = 2)
```

The above plot shows the polynomial fits to the data with the red line being the fit for a polynomal with order 1, grey having a polynomial function with order 2, black having a polynomial function with order 3, and purple having a polynomial function with order 4.

#Part C

```{r}
#Use cross validation to find degrees of freedom
fit.ss <- smooth.spline(concrete$water, concrete$compressive_strength)
fit.ss$df

pred.smooth <- predict(fit.ss, x = water.grid)

pred.smooth.df <- data.frame(pred = pred.ss$y,
                         water = water.grid)

fitplot <- ggplot(data = concrete, aes(x = water, y = compressive_strength)) +
  geom_point(color = rgb(.2, .2, .4, .1)) + 
  geom_line(aes(x = water, y = pred), data = pred.smooth.df, 
  color = rgb(.8, .1, .1, 1)) + theme_bw() 
fitplot
```

Using the cross validation method the degrees of freedom is 68.88 which we round up to 69. 

```{r}
#fitting a range of degrees of freedom including the degrees of freedom from the cross validation method

fit.ss2 <- smooth.spline(concrete$water, concrete$compressive_strength, df = 2)
fit.ss2
pred.ss2 <- predict(fit.ss2, x = water.grid)
pred.ss.df2 <- data.frame(pred = pred.ss2$y, water = water.grid)


fit.ss10 <- smooth.spline(concrete$water, concrete$compressive_strength, df = 10)
fit.ss10
pred.ss10 <- predict(fit.ss10, x = water.grid)
pred.ss.df10 <- data.frame(pred = pred.ss10$y, water = water.grid)


fit.ss50 <- smooth.spline(concrete$water, concrete$compressive_strength, df = 50)
fit.ss50
pred.ss50 <- predict(fit.ss50, x = water.grid)
pred.ss.df50 <- data.frame(pred = pred.ss50$y, water = water.grid)


fit.ss <- smooth.spline(concrete$water, concrete$compressive_strength)
fit.ss
pred.ss <- predict(fit.ss, x = water.grid)
pred.ss.df <- data.frame(pred = pred.ss$y, water = water.grid)

#plotting the fits by overlaying them 
fitting_df <- ggplot(data = concrete, aes(x = water, y = compressive_strength)) +
  geom_point(color = rgb(.2, .4, .2, .5)) + 
  geom_line(aes(x = water, y = pred), data = pred.ss.df2,
          color = rgb(.8, .1, .8, 1)) + 
  geom_line(aes(x = water, y = pred), data = pred.ss.df10,
          color = rgb(0, 0, 1, 1)) + 
  geom_line(aes(x = water, y = pred), data = pred.ss.df50,
          color = rgb(1, 0, 1, 0)) + 
  geom_line(aes(x = water, y = pred), data = pred.ss.df,
          color = rgb(.2, .2, .4, 1)) + 
  theme_bw()
fitting_df
```

As we increase the degrees of freedom, the variance in the fit model goes down. Furthermore, based on the graph as we increase the degrees of freedom, the closer the fitting line comes to connecting the mean points. These results show that with such a high degrees of freedom (i.e. 69) we could be at risk of overfitting the data. 

#Part D

```{r}
#GAM model without any smoothing spline for any variable
gam.m1 <- gam(compressive_strength ~ cement + blast_furnace_slag + fly_ash + water + superplasticizer + coarse_aggregate + fine_aggregate + age, data = concrete)

#GAM model with smoothing spline for water
gam.m2 <- gam(compressive_strength ~ cement + blast_furnace_slag + fly_ash + s(water) + superplasticizer + coarse_aggregate + fine_aggregate + age, data = concrete)

#anova test to see if the GAM models are different
anova(gam.m1, gam.m2, test = "F")
```

According to the anova test the GAM model with the smoothing spline for water is better fitting than the GAM model without the smoothing spline. 

```{r}
plot(gam.m2)
```

According to the graph above, the tails of the model have the greatest variance and the model is not very linear even with the smoothing splines. There can be futher models that we can try to get a more linear function. 